Import Libraries

```{r}
library(tidyverse)
```

Load Data

```{r}
met_objects <- read_csv("openaccess/MetObjects.csv")
met_objects
```

Look at the problems!

```{r}
problems(met_objects)
```

Cleaning it up

```{r}
met_objects$AccessionYear <- as.double(substr(met_objects$AccessionYear, 1, 4))
```

Time to analyze our columns and see which ones we should try for general predictions

```{r}
colnames(met_objects)
```

Variables that seem like they could possibly be useful for predicting `Is Highlight`

1. `Is Timeline Work`
2. `Is Public Domain`
3. `Gallery Number`
4. `Department`
5. `AccessionYear`
6. `Culture`
7. `Period`
8. `Dynasty`
9. `Reign`
10. `Portfolio`
11. `Artist Nationality`
12. `Artist Begin Date`
13. `Artist End Date`
14. `Artist Gender`
15. `Object Date`
16. `Object Begin Date`
17. `Object End Date`
18. `Medium`
19. `Dimensions`
20. `City`
21. `State`
22. `County`
23. `Country`
24. `Region`
25. `Subregion`
26. `Locale`
27. `Excavation`
28. `Classification`
29. `Rights and Reproduction`


Some of these seem useful not as direct variables, but as whether they have *any* value. These include

1. `Portfolio`
2. `Excavation`
3. `Rights and Reproduction`

So we'll have to turn these into bools.

```{r}
met_objects <- met_objects %>%
  mutate(has_portfolio = !is.na(Portfolio)) %>%
  mutate(has_excavation = !is.na(Excavation)) %>%
  mutate(has_rights = !is.na(`Rights and Reproduction`))
```

```{r}
met_objects
```


